Linear response theory for core set identification

### 2013-06-10 writedown/linear.response/paper.tex

>>>>

\section{Optimal Forcing}

Let us remain in the setting of the last section and assume that we are interested in a finite dimensional family of forcings
\[
\mathcal{F}=\{F_e(t)=\sum_{k=1}^N a_k \Phi_k(t)\},
\]
with explicitly given basis functions $\Phi_k$, $k=1,\ldots,N$.
We are interested in computing the gradient of the expectation value at time $T$ of the observable $\mathcal{O}$ wrt the coefficient vector $(a_k)_{k=1,\ldots,N}$ in the point $F_e(t)=\sum_{k=1}^N a_k \Phi_k(t)$. From Eq. (\ref{eqn:neq-response}) we get with $\Delta F_e(t)=\Phi_k(t)$:
\[
\frac{\partial O(T)}{\partial a_j} = \lim_{\eps\to 0}\frac{1}{\eps}( O_\eps(T) -
  O_0(T))=\big\langle O[\vect X_s]\cdot G_j[\vect X_s]\big\rangle_{0,T},
\]
where $X_t$ and $O_0$ are paths and expectation value wrt. the dynamics \eqref{eqn:langevin-1b}--\eqref{eqn:langevin-2b} with $F_e(t)=\sum_{k=1}^N a_k \Phi_k(t)$ and $\eps=0$, and
\[
G_j[\vect X_t]
  = \frac{1}\sigma\int_0^t
  \Phi_j(s)\,\vect D(\vect q_s)\cdot\dd\vect W_s. 
\]
That is, we can compute all derivatives $\partial O(T)/\partial a_j$ from just one ensemble of paths of $X_t$.

Now we consider the optimization problem
\[
O_{max}(T) = \max_{F_e\in \mathcal{F}} O[F_e](T),
\]
where $O[F](T)$ is the expectation value of the path observable $\mathcal{O}$ that results from \eqref{eqn:langevin-1b}--\eqref{eqn:langevin-2b} with $F_e=F$ and $\eps=0$.
With the gradient being computed one can solve the optimization problem by means of gradient descent in coefficient space,
\[
a^{(i+1)}=a^{(i)} -\gamma \frac{\partial O(T)}{\partial a}.
\]
